---
title: "Homework#1"
author: "Azadeh Gilanpour"
date: "August 30, 2018"
output: word_document
---
#Questions

##Question# 1: Using R vector
 
###(a)
create a vector with 10 specefic number and assiging to X
```{r}
x<- c(3,12,6,-5,0,8,15,1,-10,7)
x
```

###(b)
create a vector y with 10 elements of x from min to max
```{r}
y<- y <-seq(min(x),max(x),length=10)
y
```

###(c)

 * summation of elements "x" and "y" 
```{r}
sum(x); sum(y)
```
 
 * mean value of "x" and "y" 
```{r}
mean(x);mean(y)
```
 
* standard deviation values for "x" and "y"
```{r}
sd (x); sd (y) 
```

* Variances for "x" and "y"
```{r}
var(x); var(y)
```

* Mean absolute deviation values for "x" and "y"
```{r}
mad(x); mad(y)
```

* Quartiles for "x" and "y"
```{r}
quantile(x); quantile(y)
```

* Quintiles for "x" and "y"
```{r}
quantile(x, probs = seq(0, 1, 0.25)); quantile(y, probs = seq(0, 1, 0.25)) 
```

###(d)
create vector z consist 7 elemnts randomly sample from x
```{r}
z<- sample(x, 7,replace=TRUE) 
z
```

###(e) 
 package for calculaiton of skewness and kurtosis
```{r}
library(e1071)
```

* Skewness of vector "x" from package e1071
```{r}
skewness(x)
```

* kurtosis of vector "x" from package e1071
```{r}
kurtosis(x)
```


###(f)
t-test to compare the mean of "x" and "y"
```{r}
t.test(x,y)   
```

+ The reults shows the difference between two sets are not significant. The p-values for the test is %74 which is very higher than %5. The confidence interval is also (-6.32, 8.72) which contains zero.

###(g)
t-test to compare the mean of sorted "x" and "y"
```{r}
t.test(sort(x),y,paired = T)  
```

###(h)
Logical vector to show the negative number of x
```{r}
as.logical(x<0)  
```

###(i)
remove negative number from x
```{r}
x<- x[!x<0]  
x
```

##Question# 2: Using R Introductory data exploration

###(a)
Read data from file "college.csv"
```{r}
college<-read.csv("college.csv")  
```

####(b)
```{r}
# Use the first column of data as row's name
rownames(college) <- college[,1] 

# Display the content of the data frame
View (college )

# Remove the first column's of data
college <- college[,-1]          

```

###(c) 

   * i Shows the numerical summaries of "college" dataset
```{r}

summary(college)   
```

   * ii 
```{r}
#Shows the description for "pairs" function
help("pairs") 
# Used pairs fucntion to produce a scatterplot matrix for the first 10 columns
pairs(college[,1:10])
```

  * iii Side-by-side boxplot for Outstate vs. Private
```{r}
plot(Outstate~Private,data=college,
     main="Out Of state Tuitions for Private and public university",
     xlab="Private and Non-Private univeristy",
     ylab="Out Of State Tuition") 
```

   * iv
```{r}
# Creates a vector in size of rows of the college filling with No
Elite <- rep("No", nrow(college))

# Find the Top10prec greater than 50 and replace the corresponding elements of "Elite" with "Yes"
Elite [college$Top10perc>50] <- "Yes"

# the "Elite" vector into a factor vector with 2 levels
Elite <- as.factor(Elite) 

# Adds "Elite" as a new column to the "College" data frame
college <- data.frame(college, Elite) 
```

   * v Shows the number of the elite and non-elite colleges
```{r}
summary(Elite) 
```

  * vi # Side-by-side boxplot for Outstate vs. Elite
```{r}
plot(Outstate~Elite,data=college,
     main="Out of state Tuition for Elite and Non-Elite universities",
     xlab="Elite and Non-Elite University",
     ylab="Out Of State Tuition") 
```

   * vii
```{r}
# Divides the screen into 4 windows
par(mfrow=c(2,2))  
# Draw histograms for Number of new students enrolled, number of applicants accepted,Percent of faculty with Ph.D.s , and Graduation rate.
hist(college$Enroll,xlab = "Enrollment",main = "Number of student Enrollment",breaks = 5)

hist(college$Accept,xlab = "Acceptance",main = "Number of Students with N acceptance",breaks=10)

hist(college$PhD,xlab="PhD Students",main="Number of Phd students of Universities", breaks = 10)

hist(college$Grad.Rate,xlab="Gradutation Rate",main = "Rate of Student Graduation",breaks = 10)
```

##Question#3: Using R Mlating data in data frames

####(a)
```{r}
# package include of baseball dataset
library(plyr) 
#Load a baseball data set
data("baseball")
# Shows the help and description of basegall dataset from plyr package
?baseball
```

###(b)
Set 0 "sacrifies flies" for players beore 1954
```{r}
baseball$sf[baseball$year<1954]=0  
```

   + 1 Set 0 missing values for "Hit by pitch"
```{r}
baseball$hbp[is.na(baseball$hbp)]<- 0  
```
   
   + 2 Excludes all player records at bats with fewer than 50
```{r}
baseball<- subset(baseball,baseball$ab>=50)  
```
   
###(c)
```{r}
#Compute on base percentage
obp<- with(baseball, (h + bb + hbp) / (ab + bb + hbp + sf))
 #Adds "obp" as data in a new column
baseball<- data.frame(baseball,obp)
```

###(d) 
```{r}
# Sorts the sata set based on "obp"
baseball<- baseball[order(obp),] 
# Displays the top 5 on base percentage
head(baseball[,c("id", "year", "obp")], n =5)
```

##Question# 4 :Using R "aggregate()" function

###(a) 
```{r}
# package including "quakes" data set for problem #4
library(datasets) 
#Load a quakes data set
data("quakes")
```

###(b)
```{r}
# Reset display windows
par(mfrow=c(1,1))

# Plots the earthquake magnetude agianst the depth
plot(mag~depth,data=quakes,
     main= "The Earthquake magnetude vs the depth",
     xlab= "Magnetude",
     ylab = "Depth") 
```

###(c) 
Use aggregate to compute the average earthquake depth for each magnitude level. Store these results in a quakeAvgDepth
```{r}
quakeAvgDepth<- aggregate(quakes$depth,list("Magnitude"=quakes$mag),mean)
```

###(d) 
Rename the quakeAvgDepth 
```{r}
names(quakeAvgDepth) <- c("Magnitude Level","Average Depth")
```

###(e) 
Plot the magnitude vs. the average depth.
```{r}
plot( quakeAvgDepth$`Magnitude Level`~ quakeAvgDepth$`Average Depth`,
      main= "Earthquake Magnitude Level vs Average Depth",
      xlab="Average Depth",
      ylab="Magbitude Level")
```

###(f)
The mean trend showes as the average depth increases the earthquack gets weaker, 